Methods for Analyzing Models of\\ Rod-Shaped Bacteria

### 2025-09-05 paper/main.tex
mark as todo
add bullet points
remove multilayer predictions

%###################################################################################################
\subsection{Data Preparation}
\begin{enumerate}
    \item Have masks of data; reference image segmantation tools; or do it by hand
    \item ensure that colors are matching for bacteria between time-steps
    \item clean borders of images if needed
    \item know domain size; otherwise use pixels
\end{enumerate}

%###################################################################################################
\subsection{Parameter Estimation}
\label{sec:supplement-parameter-profiles}
%TODO {explain profiles}
\begin{figure}
    \centering
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-0.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-1.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-2.pdf}\\%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-3.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-4.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-5.pdf}
    \caption{Parameter estimation of growth rates for individual cells}%TODO extend caption}
    \label{fig:parameter-estimates-supplement-growth-rates}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-0.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-1.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-2.pdf}\\%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-3.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-4.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-5.pdf}%
    \caption{Parameter estimation of radii for individual cells}%TODO extend caption}
    \label{fig:parameter-estimates-supplement-radii}
\end{figure}

%###################################################################################################
\subsection{More Simulation Aspects}
\label{section:supplement-more-simulation-aspects}
%TODO insert this into the discussion later
\begin{table}[H]
    \centering
    \def\arraystretch{1.3}
    \begin{tabularx}{\textwidth}{c l X}
        &\textbf{Aspect} & \textbf{Description}\\
        \toprule
        &\textbf{(C) Cellular}\\
        \midrule
        (1) & Rod-Shaped Mechanics &
            Rod-shaped bacteria are flexible rods which are able to freely move around (off-lattice
            approach) \cite{Takeuchi2005,Ursell2014,Amir2014_2}.\\
        (2) & Growth &
            Cells grow exponentially by inserting new material either along the circular part of the
            rod or at the tip~\cite{Robert2014,Takeuchi2005}.\\
        (3) & Differentiation &
            \textit{B.subtilis} is known to differentiate into matrix-producing and
            surfactin-producing cells \cite{vanGestel2015,Lpez2010}.\\
        (4) & Division &
            The formation and continuation of van Gogh bundles is driven by cell division
            \cite{vanGestel2015}.
            Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
        (5) & Variable Parameters &
            Parameters for individual cells are not fixed values but rather taken from a
            distribution \cite{Koutsoumanis2013}.\\
        &\textbf{(CC) Cell-Cell Interactions}\\
        \midrule
        (6) & Adhesion &
            Bacteria adhere to each other at longer distances and attach when in close contact
            \cite{Verwey1947,Trejo2013}.
            The interaction between rods can be polarized \cite{Duvernoy2018}.\\
        (7) & Friction &
            Friction between cells \cite{Grant2014} can be asymmetrical \cite{Doumic2020}.\\
        &\textbf{(DC) Domain-Cell Interactions}\\
        \midrule
        (8) & External Forces &
            Bacteria tend to stick to surfaces \cite{vanLoosdrecht1989}.
            The extracellular gel exerts a force onto the cells \cite{Grant2014}.\\
        (9) & Extracellular Reactions &
            Bacteria can take up nutrients or possible secrete/take up signalling molecules
            \cite{Li2025}.\\
        \bottomrule
    \end{tabularx}
    \label{table:simulation-aspects-supplement}
    \caption{
        This table extends table~\ref{table:simulation-aspects} with additional observations.
        It has to be noted that this table includes more specific behaviour that might not be
        generalizable across all types of rod-shaped bacteria.
    }
\end{table}

%###################################################################################################
\subsection{Performance}

@@

%###################################################################################################
\subsection{Multilayer Predictions}
% TODO

\begin{itemize}
    \item Friction force against surface. perpendicular to motion; proportional to velocity.
    \item gel pressure coming from experiment (simply downwards force $F_z$)
\end{itemize}

\begin{itemize}
    \item Do we need this in this publication?
    \item Maybe this is out of scope as this is a more methodological paper.
    \item If we want to omit it, we might also opt to change the model description in the beginning
        and remove some parts of the external forces discussion.
\end{itemize}

\cite{You2019}

\begin{figure}
    \centering
    \includegraphics[width=0.49\textwidth]
        {docs/source/_static/scripts/crm_multilayer/multilayer-time-evolution.pdf}\\
    \includegraphics[width=0.49\textwidth]
        {docs/source/_static/scripts/crm_multilayer/colony-height-vs-gel_pressure.pdf}
    \includegraphics[width=0.49\textwidth]
        {docs/source/_static/scripts/crm_multilayer/colony-height-vs-time.pdf}
    \caption{TODO}
    \label{fig:multilayer}
\end{figure}

>>>>

%###################################################################################################
\subsection{Data Preparation (TODO)}
\begin{enumerate}
    \item Have masks of data; reference image segmantation tools; or do it by hand
    \item ensure that colors are matching for bacteria between time-steps
    \item clean borders of images if needed
    \item know domain size; otherwise use pixels
\end{enumerate}

%###################################################################################################
\subsection{Parameter Estimation Results (TODO)}
\label{sec:supplement-parameter-profiles}
%TODO {explain profiles}
\begin{figure}
    \centering
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-0.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-1.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-2.pdf}\\%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-3.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-4.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/growth-rate-5.pdf}
    \caption{Parameter estimation of growth rates for individual cells}%TODO extend caption}
    \label{fig:parameter-estimates-supplement-growth-rates}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-0.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-1.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-2.pdf}\\%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-3.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-4.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_fit/profiles/radius-5.pdf}%
    \caption{Parameter estimation of radii for individual cells}%TODO extend caption}
    \label{fig:parameter-estimates-supplement-radii}
\end{figure}

%###################################################################################################
\subsection{More Simulation Aspects (TODO)}
\label{section:supplement-more-simulation-aspects}
%TODO insert this into the discussion later
\begin{table}[H]
    \centering
    \def\arraystretch{1.3}
    \begin{tabularx}{\textwidth}{c l X}
        &\textbf{Aspect} & \textbf{Description}\\
        \toprule
        &\textbf{(C) Cellular}\\
        \midrule
        (1) & Rod-Shaped Mechanics &
            Rod-shaped bacteria are flexible rods which are able to freely move around (off-lattice
            approach) \cite{Takeuchi2005,Ursell2014,Amir2014_2}.\\
        (2) & Growth &
            Cells grow exponentially by inserting new material either along the circular part of the
            rod or at the tip~\cite{Robert2014,Takeuchi2005}.\\
        (3) & Differentiation &
            \textit{B.subtilis} is known to differentiate into matrix-producing and
            surfactin-producing cells \cite{vanGestel2015,Lpez2010}.\\
        (4) & Division &
            The formation and continuation of van Gogh bundles is driven by cell division
            \cite{vanGestel2015}.
            Bacteria can form multilayers during their growth phase \cite{Duvernoy2018}.\\
        (5) & Variable Parameters &
            Parameters for individual cells are not fixed values but rather taken from a
            distribution \cite{Koutsoumanis2013}.\\
        &\textbf{(CC) Cell-Cell Interactions}\\
        \midrule
        (6) & Adhesion &
            Bacteria adhere to each other at longer distances and attach when in close contact
            \cite{Verwey1947,Trejo2013}.
            The interaction between rods can be polarized \cite{Duvernoy2018}.\\
        (7) & Friction &
            Friction between cells \cite{Grant2014} can be asymmetrical \cite{Doumic2020}.\\
        &\textbf{(DC) Domain-Cell Interactions}\\
        \midrule
        (8) & External Forces &
            Bacteria tend to stick to surfaces \cite{vanLoosdrecht1989}.
            The extracellular gel exerts a force onto the cells \cite{Grant2014}.\\
        (9) & Extracellular Reactions &
            Bacteria can take up nutrients or possible secrete/take up signalling molecules
            \cite{Li2025}.\\
        \bottomrule
    \end{tabularx}
    \label{table:simulation-aspects-supplement}
    \caption{
        This table extends table~\ref{table:simulation-aspects} with additional observations.
        It has to be noted that this table includes more specific behaviour that might not be
        generalizable across all types of rod-shaped bacteria.
    }
\end{table}

\begin{itemize}
    \item new class of aspects: (DC) Domain-Cell Interactions
    \item List which aspects are here additionally
    \item discuss which of them might need specific models and can not be paired with others
    \item which can be bundled together without problems?
\end{itemize}

%###################################################################################################
\subsection{Performance (TODO)}

@@

%###################################################################################################
% \subsection{Multilayer Predictions}
% % TODO
%
% \begin{itemize}
%     \item Friction force against surface. perpendicular to motion; proportional to velocity.
%     \item gel pressure coming from experiment (simply downwards force $F_z$)
% \end{itemize}
%
% \begin{itemize}
%     \item Do we need this in this publication?
%     \item Maybe this is out of scope as this is a more methodological paper.
%     \item If we want to omit it, we might also opt to change the model description in the beginning
%         and remove some parts of the external forces discussion.
% \end{itemize}
%
% \cite{You2019}
%
% \begin{figure}
%     \centering
%     \includegraphics[width=0.49\textwidth]
%         {docs/source/_static/scripts/crm_multilayer/multilayer-time-evolution.pdf}\\
%     \includegraphics[width=0.49\textwidth]
%         {docs/source/_static/scripts/crm_multilayer/colony-height-vs-gel_pressure.pdf}
%     \includegraphics[width=0.49\textwidth]
%         {docs/source/_static/scripts/crm_multilayer/colony-height-vs-time.pdf}
%     \caption{TODO}
%     \label{fig:multilayer}
% \end{figure}
